TreehopperFitness: Run this file for all stats, tables, and figures

### 2023-07-28 manuscript/Manuscript_runfile.Rmd
init `Manuscript_runfile.Rmd`; see issue #2

strats_data_with_filters <- strats_data %>%
  # Add whether female was visited within first 4 days of initiating
  group_by(location) %>%
    mutate(visited_1st_4d = any(!mom_at_init & as.numeric(visit_starts-initiation_date)<=4)) %>%
  ungroup() %>%
  # Add some additional metrics for human readability
  mutate(`visited in 4 days` = ifelse(visited_1st_4d | num_at_init>1,
                                      "Visited", "Not visited"),
         ants = ifelse(ants_removed, "Ants removed", "Control"),
         est_hatch = as_date(ifelse(is.na(hatch_date),
                                 initiation_date + 20,
                                 hatch_date)),
         est_rel2h = visit_ends-est_hatch) %>%
  # Add day of first visit bins
  left_join(first_visit_cuts) %>%
  mutate(vis_cat = ifelse(num_at_init > 1,
                          "0",
                          ifelse(is.na(first_visit_starts),
                                 "never visited",
                                 ifelse(first_visit_cut=="(0,4]", "1 to 4", "5+"))),
         vis_cat = factor(vis_cat, levels = c("0","1 to 4",
                                              "5+","never visited"))) %>%
  # Remove one female who has dubious data (See note ** below)
  filter(!is.na(vis_cat)) %>%
  # Build logicals representing inclusion in different plots
  mutate(`Text: 4 different strats` =
           !females_removed & 
           clutch_visited_order == 1,
         `Figure 4 (dur vs 1st visit)` = 
           !females_removed & 
           mom_at_init &
           clutch_initiated_order == 1 & 
           !grepl("U", Female_ID) & 
           visit_ends < last_day_fID,
         `Figure 3 (dur histograms)` = 
           !females_removed & 
           mom_at_init &
           clutch_initiated_order == 1 &
           !grepl("U", Female_ID) & 
           visit_ends < last_day_fID & 
           !is.na(`visited in 4 days`))

# ** There was one female (E_Gkgg) which was on the eggs before and after but not ON the initiation date. I suspect I missed her in the field or she may have left and come back. In any case she was never visited and deserted 20 days after initiation and we have plenty such data points, so I think it's safe to ignore.
```

*To double-check that we understand how we've filtered the data for each sub-analysis, we can represent each filtering regime as an upset plot*
```{r, fig.width=6, fig.height=2}
sample_size_upset_data <-strats_data_with_filters %>%
  mutate(`ants excluded` = ants_removed,
         `female removed` = females_removed,
         `female present at intiation` = mom_at_init,
         `her first clutch` = clutch_visited_order == 1,
         `first clutch she initiated` = clutch_initiated_order == 1,
         `female marked` = !grepl("U", Female_ID),
         `female disappeared` = !(visit_ends < last_day_fID),
         `unknown whether visited` = is.na(`visited in 4 days`)) %>%
  select(Female_ID, location,
         `ants excluded`, `female removed`,
         `female present at intiation`, 
         `first clutch she initiated`,
         `her first clutch`,
         `female marked`, `female disappeared`, `unknown whether visited`, 
         `Figure 3 (dur histograms)`, 
         `Figure 4 (dur vs 1st visit)`,
         `Text: 4 different strats`) %>%
  pivot_longer(cols = `ants excluded`:`unknown whether visited`,
               names_to = "filter_name",
               values_to = "filter_true") %>%
  filter(filter_true) %>%
  group_by(Female_ID, location, 
           `Text: 4 different strats`,
           `Figure 3 (dur histograms)`, `Figure 4 (dur vs 1st visit)`) %>%
  summarize(filters = list(filter_name))

sample_size_upset_data %>% ggplot(aes(x=filters, fill = `Text: 4 different strats`)) +
    geom_bar() +
    scale_x_upset() +
  theme_publilia()

sample_size_upset_data %>% ggplot(aes(x=filters, fill = `Figure 3 (dur histograms)`)) +
    geom_bar() +
    scale_x_upset() +
  theme_publilia()

sample_size_upset_data %>% ggplot(aes(x=filters, fill = `Figure 4 (dur vs 1st visit)`)) +
    geom_bar() +
    scale_x_upset() +
  theme_publilia()
```

## Text 
Most females laid their first eggs by initiating their own clutch while some females started by visiting an existing clutch (n=115 and 36 respectively). 
```{r}
strats_data_with_filters %>%
  filter(`Text: 4 different strats`) %>%
  count(clutch_initiated_order == 1)

visited_1st_4d_data_tb %>%
  filter(!females_removed,
         clutch_visited_order == 1) %>%
  count(clutch_initiated_order == 1)

print(paste("% (not removed) females that started the season by visiting an existing clutch:", 36/151*100))
print(paste("% (not removed) females that started by initiating their own clutch:", 115/151*100))

visited_1st_4d_data_tb %>%
  filter(females_removed,
         clutch_visited_order == 1) %>%
  count(clutch_initiated_order == 1)
62/100

visited_1st_4d_data_tb %>%
  filter(clutch_visited_order == 1) %>%
  count(clutch_initiated_order == 1)
62/215
```

>>>>

strats_data_with_filters <- strats_data %>%
  # Add whether female was visited within first 4 days of initiating
  group_by(location) %>%
    mutate(visited_1st_4d = any(!mom_at_init & as.numeric(visit_starts-initiation_date)<=4)) %>%
  ungroup() %>%
  # Add some additional metrics for human readability
  mutate(`visited in 4 days` = ifelse(visited_1st_4d | num_at_init>1,
                                      "Visited", "Not visited"),
         ants = ifelse(ants_removed, "Ants removed", "Control"),
         est_hatch = as_date(ifelse(is.na(hatch_date),
                                 initiation_date + 20,
                                 hatch_date)),
         est_rel2h = visit_ends-est_hatch) %>%
  # Add day of first visit bins
  left_join(first_visit_cuts) %>%
  mutate(vis_cat = ifelse(num_at_init > 1,
                          "0",
                          ifelse(is.na(first_visit_starts),
                                 "never visited",
                                 ifelse(first_visit_cut=="(0,4]", "1 to 4", "5+"))),
         vis_cat = factor(vis_cat, levels = c("0","1 to 4",
                                              "5+","never visited"))) %>%
  # Remove one female who has dubious data (See note ** below)
  #filter(!is.na(vis_cat)) %>%
  # Build logicals representing inclusion in different plots
  mutate(`Text: 4 different strats` =
           !females_removed & 
           clutch_visited_order == 1 &
           !is.na(num_at_init),
         `Figure 4 (dur vs 1st visit)` = 
           !females_removed & 
           mom_at_init &
           clutch_initiated_order == 1 & 
           !grepl("U", Female_ID) & 
           visit_ends < last_day_fID &
           !is.na(num_at_init),
         `Figure 3 (dur histograms)` = 
           !females_removed & 
           mom_at_init &
           clutch_initiated_order == 1 &
           !grepl("U", Female_ID) & 
           visit_ends < last_day_fID & 
           !is.na(`visited in 4 days`) &
           !is.na(num_at_init))

# ** There was one female (E_Gkgg) which was on the eggs before and after but not ON the initiation date. I suspect I missed her in the field or she may have left and come back. In any case she was never visited and deserted 20 days after initiation and we have plenty such data points, so I think it's safe to ignore.

# ** 2023.07.28: c("D_Rwyw", "E_Rgrw", "E_Gkgg", "F_Yrgw") on c("D82_12", "E148_6", "E40_9",  "F4_6") were similar. See github issue #2.
```

*To double-check that we understand how we've filtered the data for each sub-analysis, we can represent each filtering regime as an upset plot*
```{r, fig.width=10, fig.height=4}
sample_size_upset_data <- strats_data_with_filters %>%
  mutate(`ants excluded` = ants_removed,
         `female removed` = females_removed,
         `female present at intiation` = mom_at_init,
         `her first clutch` = clutch_visited_order == 1,
         `first clutch she initiated` = clutch_initiated_order == 1,
         `female marked` = !grepl("U", Female_ID),
         `female disappeared` = !(visit_ends < last_day_fID),
         `unknown whether visited` = is.na(`visited in 4 days`),
         `initiator known` = !is.na(num_at_init)) %>%
  select(Female_ID, location,
         `ants excluded`, `female removed`,
         `female present at intiation`, 
         `first clutch she initiated`,
         `her first clutch`,
         `female marked`, `female disappeared`, `unknown whether visited`, `initiator known`, 
         `Figure 3 (dur histograms)`, 
         `Figure 4 (dur vs 1st visit)`,
         `Text: 4 different strats`) %>%
  pivot_longer(cols = `ants excluded`:`initiator known`,
               names_to = "filter_name",
               values_to = "filter_true") %>%
  filter(filter_true) %>%
  group_by(Female_ID, location, 
           `Text: 4 different strats`,
           `Figure 3 (dur histograms)`, `Figure 4 (dur vs 1st visit)`) %>%
  summarize(filters = list(filter_name))

sample_size_upset_data %>% 
  ggplot(aes(x=filters, fill = `Text: 4 different strats`)) +
    geom_bar() +
    scale_x_upset() +
    geom_text(stat='count', aes(label=after_stat(count)), vjust=-1) +
  theme_publilia()

sample_size_upset_data %>% 
  ggplot(aes(x=filters, fill = `Figure 3 (dur histograms)`)) +
    geom_bar() +
    scale_x_upset() +
  geom_text(stat='count', aes(label=after_stat(count)), vjust=-1) +
  theme_publilia()

sample_size_upset_data %>% 
  ggplot(aes(x=filters, fill = `Figure 4 (dur vs 1st visit)`)) +
    geom_bar() +
    scale_x_upset() +
  geom_text(stat='count', aes(label=after_stat(count)), vjust=-1) +
  theme_publilia()
```

## Text 
```{r}
n_start_by_init <- strats_data_with_filters %>%
  filter(`Text: 4 different strats`) %>% 
  pull(clutch_initiated_order) %>% 
  sum(.data == 1, na.rm = TRUE)

n_start_by_visit <- strats_data_with_filters %>%
  filter(`Text: 4 different strats`) %>% 
  pull(clutch_initiated_order) %>% 
  is.na() %>% sum()

# visited_1st_4d_data_tb %>%
#   filter(!females_removed,
#          clutch_visited_order == 1) %>%
#   count(clutch_initiated_order == 1)

# print(paste("% (not removed) females that started the season by visiting an existing clutch:", 36/151*100))
# print(paste("% (not removed) females that started by initiating their own clutch:", 115/151*100))
# 
# visited_1st_4d_data_tb %>%
#   filter(females_removed,
#          clutch_visited_order == 1) %>%
#   count(clutch_initiated_order == 1)
# 62/100
# 
# visited_1st_4d_data_tb %>%
#   filter(clutch_visited_order == 1) %>%
#   count(clutch_initiated_order == 1)
# 62/215
```
Most females laid their first eggs by initiating their own clutch while some females started by visiting an existing clutch (n=`r n_start_by_init` and `r n_start_by_visit` respectively).
